RL-Musician: A Tool for Music Composition with Deep Reinforcement Learning

### 2019-10-12 docs/paper/paper.tex
docs: Update some paragraphs of paper

Brief outline of the approach is as follows. There is an environment with 2D table representing piano roll which means that rows of the table correspond to consecutive notes and columns of the table correspond to time steps. An agent observes current time step and some previous time steps. An action of the agent can be either playing a note at the current time step or moving to the next time step. When the last time step of the piano roll is reached, episode is ended and the agent receives reward based on evaluation of created composition. Trainable parameters of the agent are weights of a neural network used as a so called actor model, i.e., a model that maps observation to probabilities of actions. These parameters are trained with cross-entropy method \cite{rubinstein1997optimization} (alternatively, they can be trained with genetic algorithms, ant colony algorithms, or evolutionary strategies \cite{salimans2017evolution}).

@@

Further, sequential generation of composition by neural network implies that any well-studied architectures from sequence generation tasks can be plugged in and tested. The only difference is that now these neural networks can not be trained with variations of gradient descent, but can be trained with cross-entropy method or alternative methods.

@@

Nevertheless, not all methods for automated music generation use deep learning. Musical pieces itself can be generated with genetic algorithms \cite{felice2002genorchestra} where selection is based on scores returned by hand-written evaluation rules.

@@

>>>>

Brief outline of the approach is as follows. There is an environment with 2D table representing piano roll which means that rows of the table correspond to notes and columns of the table correspond to time steps. An agent observes current time step and some previous time steps. An action of the agent is playing a note at the current time step. After certain number of actions are done, a movement to the next time step happens. When the last time step of the piano roll is reached, episode is ended and the agent receives reward based on evaluation of created composition. Trainable parameters of the agent are weights of a neural network used as a so called actor model, i.e., a model that maps observation to probabilities of actions. These parameters are trained with cross-entropy method \cite{rubinstein1997optimization} (alternatively, they can be trained with genetic algorithms, ant colony algorithms, or evolutionary strategies \cite{salimans2017evolution}).

@@

@@

Nevertheless, not all methods for automated music generation use deep learning. Musical pieces itself can be generated with genetic algorithms \cite{felice2002genorchestra} or Markov chains \cite{farbood2001analysis}. Comprehensive review that covers plenty of approaches beyond deep learning can be found in \cite{fernandez2013ai}.

@@

In context of reinforcement learning, cross-entropy method is used for searching parameters of agent that result in maximum expected reward. Such approach is best suitable for environments where reward becomes known only after episode is finished. If there are rewards granted after intermediate steps, methods based on Q-values usually outperform it.
